Study of the hadronization of charged pions

### 2017-03-23 chap5.tex
Small fixes (pt max = 1.5 GeV2)

Once the identification is done, the calculation of our two observables and their statistical uncertainties is straightforward. The latter are given by:
\begin{equation}
\label{eq:MRError}
{\delta \left ( R_A^h \right ) \over R_A^h} = \sqrt{ 1 /N_A^h + 1/ N_A^e +1/N_D^h + 1/N_D^e}
\end{equation}
and
\begin{equation}
\label{eq:DPtError}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_A / N_A^h
 + \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_D / N_D^h.
\end{equation}

@@

Following the first round of questions from the analysis review committee, we
reflected about this implementation of error bars and possible overlap with the 
systematic errors
evaluated below. We found first that the correlation between the weights should reduce
the statistical errors. In a given bin the weights associated with each 
targets are always of similar sizes, this correlation will reduce the size of 
error bars. However, we do not see a method to properly take this 
effect fully into account in term of purely statistical error bars. In order
to solve this issue we came back to a point by the committee on the topic of 
systematic error bars (Q 24). There it was noted that some of these effects are
double counted once in the statistical error and once in the systematic. As these
errors are solely due to the acceptance correction, we decided to account them 
only in the systematic error budget. So we reverted statistical errors to equations
\ref{eq:MRError} and \ref{eq:DPtError} and keep the full systematic errors as they
were evaluated before.

@@

\paragraph{Simulation} ~\\
To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process in a fashion similar 
to that done with the experimental data.
% TODO GPP factors

@@

\paragraph{Correction of the Data} ~\\
The acceptance correction coefficients are defined as the ratio of reconstructed and generated events,

@@

\begin{table}[htbp]
\center
\begin{tabular} {|c|c|c|c|c|}
\hline
 $\nu$ & \xb & $p_\pi$ & $-t$ \\ \hline
 2.20 & 0.12 & 0.25 &-3.50 \\ 
 3.00 & 0.18 & 0.60 &-2.10 \\ 
 3.40 & 0.22 & 1.00 &-1.75 \\ 
 3.75 & 0.27 & 1.45 &-1.50 \\ 
 4.00 & 0.33 & 1.70 &-1.30 \\ 
 4.25 & 0.60 & 2.15 &-1.10 \\ 
      &      & 3.00 &-0.75 \\ 
      &      & 4.25 &-0.00 \\ \hline
\end{tabular}
\quad
\begin{tabular} {|c|c|c|c|c|}
\hline
 $Q^2$ & $\nu$ & $z$ & $-t$ \\ \hline
 1.00 & 2.20 & 0.00 & -3.50 \\
 1.20 & 3.00 & 0.20 & -2.10 \\
 1.40 & 3.40 & 0.40 & -1.75 \\
 1.65 & 3.75 & 0.45 & -1.50 \\
 2.10 & 4.00 & 0.52 & -1.30 \\
 4.00 & 4.25 & 0.62 & -1.10 \\
      &      & 0.80 & -0.75 \\
      &      & 1.00 & -0.00 \\ \hline
\end{tabular}
\caption{Bin limits used for the acceptance correction described in table \ref{tab:AcceptBinning}.}
  \label{tab:AcceptBins}
\end{table}

@@

The large electric charge of our nuclear targets create an electric field that 
should be taken into account at our energies. Besides the fact that the nuclear 
electric field accelerates or decelerates the incoming and outgoing charged 
particles, it has some non trivial effects such as the focusing of low energy 
particles. Aste {\it et al.}~\cite{Aste:2005wc} showed that these Coulomb 
effects could be corrected for by using an effective momentum approximation at 
large momenta ({\it i.e.}, $Q^2>.5$ GeV$^2$). As our kinematics coverage is 
beyond this limit, we decided to use this approximation on the Coulomb 
correction of our data-sets. According to Aste {\it et al.}, the best effective 
potential approximation is given by $\bar V= 0.8 V_0$, where $V_0$ is the 
electrostatic potential at the center of nuclei. The latter is evaluated as 
$V_0= -3 \alpha_{EM} Z / (2 R)$, where $\alpha_{EM}$ is the fine-structure 
constant, $Z$ is the proton number, and R is the radius of nuclei. Table 
\ref{tab:Coulomb} contains the used values of $V_0$ and $\bar V$ on different 
nuclear target's analysis. The Coulomb correction is applied to both electrons 
and charged pions before calculating the plotted kinematical variables.

@@

The QED diagrams shown in figure \ref{fig:FDRadCorr} contribute to the measurement
of the leptoproduction of hadrons. The diagram (a) is the Born cross section, the
others are radiative contributions that we would like to subtract in order to 
have results more easily interpretable in term of hadronic physics.
This substraction is what we call the radiative correction, as can be seen 
this correction should take care of real photon productions (diagrams b and c)
but also virtual photon exchange (diagram d) and photon self energy (diagram e).
We neglect exchanges of photons with hadrons as they are suppressed due to the 
large mass of hadrons in comparison to electrons. For practical reasons, we also 
limit this correction to leading order as can be seen from the presented diagrams.

\begin{figure}[htp]
\centering
\includegraphics[width=12cm] {chap5-fig/RadDiag.png}
\caption {Diagrams taken into account in the HAPRAD 2 code \cite{Akushevich:2007jc}. The 
diagrams b to e contribute to the first order calculation of the radiative 
effects of the Born cross section represented in diagram a.}
\label{fig:FDRadCorr}
\end{figure}

The radiative correction is not expected to be important in the multiplicity 
ratios due to the double cancellation between the different contributions of 
the ratio. However, its impact on the transverse momentum brodening is more 
difficult to anticipate. We therefore decided to calculate the correction using 
the recent Haprad 2 code \cite{Akushevich:2007jc} with adaptations to include
nuclear effects such as Fermi motion \cite{Osipenko:PC}. This calculation is
using modeled semi-inclusive structure functions (based on former CLAS
data \cite{Akushevich:2007jc}) and modeled exclusive channels from MAID 2003
\cite{Drechsel:1998hk}. This calculation is the state of the art in the matter
as it is an exact calculation in the complete five dimension cross section
and do not use the peaking approximation generally used in previous method.

To apply the radiative correction, we apply a weight to each events in a fashion
similar to the acceptance correction. The HAPRAD 2 code provides 
$\delta = \sigma_{obs} / \sigma_B$ at given kinematics. Table \ref{tab:RCBins} 
shows the binning used to calculate the radiative correction weights.
We chose to make a finer 
binning for the radiative corrections than for acceptance as the computing time 
to calculate weights is more reasonnable (around 1000 days of cpu time).

\begin{table}[htbp]
\center
\begin{tabular} {|c|c|c|c|c|c|c|}
\hline
Bin Nb & $\nu$ & \xb & $p_\pi$ & $-t$ & $\phi$ \\ \hline

 0     & 2.25 & 0.12 & 0.20 & -3.50 &   0.0 \\ 
 1     & 2.45 & 0.15 & 0.35 & -2.10 &  18.0 \\ 
 2     & 2.66 & 0.17 & 0.50 & -1.75 &  36.0 \\ 
 3     & 2.82 & 0.18 & 0.75 & -1.50 &  54.0 \\ 
 4     & 2.96 & 0.19 & 1.00 & -1.30 &  72.0 \\ 
 5     & 3.07 & 0.20 & 1.60 & -1.10 &  90.0 \\ 
 6     & 3.17 & 0.21 & 2.20 & -0.75 & 108.0 \\ 
 7     & 3.27 & 0.22 & 4.25 &  0.00 & 126.0 \\ 
 8     & 3.37 & 0.23 &      &       & 144.0 \\ 
 9     & 3.46 & 0.24 &      &       & 162.0 \\ 
10     & 3.55 & 0.25 &      &       & 180.0 \\ 
11     & 3.64 & 0.26 &      &       &       \\ 
12     & 3.72 & 0.27 &      &       &       \\ 
13     & 3.80 & 0.28 &      &       &       \\ 
14     & 3.87 & 0.30 &      &       &       \\ 
15     & 3.94 & 0.33 &      &       &       \\ 
16     & 4.01 & 0.36 &      &       &       \\ 
17     & 4.07 & 0.39 &      &       &       \\ 
18     & 4.12 & 0.43 &      &       &       \\ 
19     & 4.18 & 0.48 &      &       &       \\ 
20     & 4.25 & 0.56 &      &       &       \\

\hline
\end{tabular}
\caption{Bin limits used for the radiative correction described in section \ref{RadCor}.}
  \label{tab:RCBins}
\end{table}

We show in figure \ref{fig:RCexample} a selected sample of radiative correction factors
($= \delta_{Fe} / \delta_D$) in order to illustrate the expected effect on observables.
In the large majority of the cases the factors are within 5\% of 1, so we do not expect to
see a strong effect on the final results where these will be integrated. This is confirmed 
in figure \ref{} ... In the case of the transverse momentum braodening, the variations
appear to ...

\begin{figure}[tbp]
\centering
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_2356.png}
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_3756.png}
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_7756.png}
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_10756.png}
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_11556.png}
\includegraphics[width=7cm] {chap5-fig/ElecRadWei_Iron_18956.png}
\caption {RC factors ($= \delta_{Fe} / \delta_D$) as a function of $\phi$ and 
$-t$ for a selection of bins specified above each figure.}
\label{fig:IsoSpin}
\end{figure}

\begin{figure}[tbp]
\centering
\caption {Preliminary results ..., 
for multiplicity ratios (top) and transverse momentum broadening 
(bottom). Only statistical errors are shown.}
\label{fig:RCPlots}
\end{figure}

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=10cm] {chap5-fig/HallC-Isospin.png}
\caption {}
\label{fig:RCexample}
\end{figure}

@@

We also evaluated the particles' misidentification and scattering on materials
from the detectors or coils. For this study, we are looking at reconstructed 
particles that do not have a clear parent (either because they are misidentified
or because they scattered and their kinematic changed significantly). The 
misidentification of electrons was found to be small, of the order of 1 in 
1000 $e^-$ in all the kinematic space, hence this issue does not contribute 
significantly to this uncertainty. We found however for pions
a high proportion of accidental events on the kinematical distributions tails, 
in particular of the \pt distribution. This effect is due to the low 
probability to produce high \pt events, leading to an increased relative 
contribution from the accidental background. As a consequence, we used the 
following cut $p_\perp^2 < 2.5$~GeV$^2$/c$^2$, which discards only a very 
little amount of data (about 1 in 30000 $\pi$). On the contamination side,
most events come from kaons with momenta above 2~GeV/c ($\sim$ 3\% for 
$\pi^+$s and $\sim$ 0.5\% for $\pi^-$s). Protons also contaminate the 
$\pi^+$ sample but only for the highest momenta (up to a few \%).

>>>>

Once the identification is done, the calculation of our two observables and their statistical uncertainties is straightforward. The latter are given by:
\begin{equation}
{\delta \left ( R_A^h \right ) \over R_A^h} = \sqrt{ 1 /N_A^h + 1/ N_A^e +1/N_D^h + 1/N_D^e}
\end{equation}
and
\begin{equation}
\left ( \delta \left ( \Delta \langle P_\perp^2 \rangle \right ) \right )^2 = 
   \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_A / N_A^h
 + \left ({\langle P_\perp^4 \rangle - \langle P_\perp^2 \rangle ^2}\right )_D / N_D^h.
\end{equation}

@@

@@

\paragraph{Simulation}
\label{sec:simul}

To correct for acceptance effects, we simulated a total of 100 million events 
per target ($^2$H, C, Fe and Pb) using the PYTHIA \cite{Sjostrand:2006za} 
event generator that was slightly modified to include Fermi motion effects. The 
generated events are processed by CLAS softwares (GSIM, GPP and user\_ana) 
to simulate the detector and the reconstruction process in a fashion similar 
to that done with the experimental data.
% TODO GPP factors

@@

\paragraph{Correction of the Data}

The acceptance correction coefficients are defined as the ratio of reconstructed and generated events,

@@

\begin{table}[htbp]
\center
\begin{tabular} {|l|c|c|c|c|}
\hline
 $\nu$ & \xb & $p_\pi$ & $-t$ \\ \hline
 2.20 & 0.12 & 0.25 &-3.50 \\ 
 3.00 & 0.18 & 0.60 &-2.10 \\ 
 3.40 & 0.22 & 1.00 &-1.75 \\ 
 3.75 & 0.27 & 1.45 &-1.50 \\ 
 4.00 & 0.33 & 1.70 &-1.30 \\ 
 4.25 & 0.60 & 2.15 &-1.10 \\ 
      &      & 3.00 &-0.75 \\ 
      &      & 4.25 &-0.00 \\ \hline
\end{tabular}
\quad
\begin{tabular} {|l|c|c|c|c|}
\hline
 $Q^2$ & $\nu$ & $z$ & $-t$ \\ \hline
 1.00 & 2.20 & 0.00 & -3.50 \\
 1.20 & 3.00 & 0.20 & -2.10 \\
 1.40 & 3.40 & 0.40 & -1.75 \\
 1.65 & 3.75 & 0.45 & -1.50 \\
 2.10 & 4.00 & 0.52 & -1.30 \\
 4.00 & 4.25 & 0.62 & -1.10 \\
      &      & 0.80 & -0.75 \\
      &      & 1.00 & -0.00 \\ \hline
\end{tabular}
\caption{Bin limits used for the acceptance correction described in table \ref{tab:AcceptBinning}.}
  \label{tab:AcceptBins}
\end{table}

@@

The large electric charge of our nuclear targets create an electric field that should be taken into account at our energies. Besides the fact that the nuclear electric field accelerates or decelerates the incoming and outgoing charged particles, it has some non trivial effects such as the focusing of low energy particles. Aste {\it et al.}~\cite{Aste:2005wc} showed that these Coulomb effects could be corrected for by using an effective momentum approximation at large momenta ({\it i.e.}, $Q^2>.5$ GeV$^2$). As our kinematics coverage is beyond this limit, we decided to use this approximation on the coulomb correction of our data-sets. According to Aste {\it et al.}, the best effective potential approximation is given by $\bar V= 0.8 V_0$, where $V_0$ is the electrostatic potential at the center of nuclei. The latter is evaluated as $V_0= -3 \alpha_{EM} Z / (2 R)$, where $\alpha_{EM}$ is the fine-structure constant, Z is the proton number, and R is the radius of nuclei. Table 5 contains the used values of $V_0$ and $\bar V$ on different nuclear target's analysis. We must note that this Coulomb correction has been applied to both electrons and charged pions before calculating any kinematical variables.

@@

Despite its lower magnitude compared to the strong force, the 
electromagnetic force may have a non negligible impact on our results. The 
reason is that even a moderate energy photon emission can modify significantly the measured kinematical variables. To correct this effect, several simulation codes exist \cite{Akushevich:2001yp}, however, none of them treat directly the SIDIS production on nuclei. For that reason, a dedicated Monte-Carlo simulation based on existing codes was developed.

\paragraph{Simulation}

The inclusive radiative effects are generated using the RADGEN code 
\cite{Akushevich:1998ft}, which includes Feynman diagrams shown in 
figure~\ref{fig:FDRadCorr}. Using this code one can evaluate the correction 
factors for inclusive measurements as $\delta = \sigma_{obs} / \sigma_{Born}$, where $\sigma_{obs}$ is the measured inclusive cross section, and $\sigma_{Born}$ is the Born cross section. As RADGEN is also a Monte-Carlo generator for photon emissions, it could be used to modify the virtual photon kinematics before the 
hadron production in any DIS generator. This allows the evaluation of the radiative 
effects on semi-inclusive measurements by implementing RADGEN inside the main 
Monte-Carlo event generator.

\begin{figure}[htp]
\centering
\includegraphics[width=12cm] {chap5-fig/RadDiag.png}
\caption {Diagrams taken into account in RADGEN \cite{Akushevich:1998ft}. The 
diagrams b) to e) contribute to the first order calculation of the radiative 
effects on Born cross section (diagram a)).}
\label{fig:FDRadCorr}
\end{figure}

The Monte-Carlo simulation was done using the event generator, GENIE~\cite{Andreopoulos:2009rq}, that describes well electron-nucleus interactions. Indeed, GENIE contains the quasi-elastic scattering\footnote{Elastic 
scattering off a single nucleon on nuclei.}, which is considered as the main channel generating radiative events in the DIS region \cite{Akushevich:2007jc}. This generator includes also hadronic cascades in nuclei, at which pions might be generated in quasi-elastic events. This is an important feature because the quasi-elastic production can not contribute directly to semi-inclusive measurements.

\paragraph{Correction of the Data}

We generated 100 million events with and without radiative effects for each target. The RADGEN code gave us the magnitude of inclusive correction coefficients, however, the comparison between the two data sets allows the extraction of semi-inclusive correction factors.

The ratio of inclusive correction factors, as depicted in figure \ref{fig:RadCorrFac}, is relatively small (couple of percents at most) except for low \xb and high $\nu$. Similarly for the semi-inclusive factors' ratio shown in figure \ref{fig:RadCorrFacSIDIS} as a function of $z$ and \ptp. However, the associated statistical errors are quite large because only a small fraction of the events involved a photon radiation.

\begin{figure}[htbp]
\centering
\includegraphics[width=9cm] {chap5-fig/ElecRadWei_Lead.png}
\caption {Ratios of radiative correction factors, $\delta_\text{Pb}/
\delta_{^2\text{H}}$, as a function of \xb in various $\nu$ bins.}
\label{fig:RadCorrFac}
\end{figure}

\begin{figure}[htbp]
\centering
\includegraphics[width=9cm] {chap5-fig/RadGenFactors.png}
\caption {Ratios of radiative correction factors, $\delta_\text{Pb}/
\delta_{^2\text{H}}$, as a function of $z$ (top) and \ptp (bottom).}
\label{fig:RadCorrFacSIDIS}
\end{figure}

The semi-inclusive part of the radiative correction turned out to be limited 
in amplitude for the most kinematics, except for high $z$ and \ptp. Thus, we decided to not apply this correction before further studies.

{\color{red} We found a better way to implement the radiative correction,
that is not limited by the used MC technique. We will implement it in
the second version of the note as we do not want to delay the review process because this correction that we already found to be small. Sorry about that.}

@@

\begin{figure}[tbp]
\centering
\includegraphics[width=10cm] {chap5-fig/HallC-Isospin.png}
\caption {Ratios of deuteron over proton for $\pi^+$s (top) and $\pi^-$s (bottom)
as a function of \xb at $z=0.55$ \cite{Asaturyan:2011mq}.}
\label{fig:IsoSpin}
\end{figure}

@@

We also evaluated the particles' misidentification and scattering on materials
from the detectors or coils. For this study, we are looking at reconstructed 
particles that do not have a clear parent (either because they are misidentified
or because they scattered and their kinematic changed significantly). The 
misidentification of electrons was found to be small, of the order of 1 in 
1000 $e^-$ in all the kinematic space, hence this issue does not contribute 
significantly to this uncertainty. We found however for pions
a high proportion of accidental events on the kinematical distributions tails, 
in particular of the \pt distribution. This effect is due to the low 
probability to produce high \pt events, leading to an increased relative 
contribution from the accidental background. As a consequence, we used the 
following cut $p_\perp^2 < 1.5$~GeV$^2$/c$^2$, which discards only a very 
little amount of data (about 1 in 30000 $\pi$). On the contamination side,
most events come from kaons with momenta above 2~GeV/c ($\sim$ 3\% for 
$\pi^+$s and $\sim$ 0.5\% for $\pi^-$s). Protons also contaminate the 
$\pi^+$ sample but only for the highest momenta (up to a few \%).
